% ---------------------------------------------------------------------------- %

\section{Problem Statement}
\label{sec:problem}

The objective of the assignment to which this report pertains is to develop a distributed messaging system with global message ordering and persistence.
The system is composed by an arbitrary number of servers, and clients are allowed to connect to any server by choice. A client, when connected to a server performs operations impacting the overall state of the system. % In order to achieve persistency, this state should be the same


\paragraph{Requirements}

These are the properties proposed in the assignment that the system implementation should abide:

\begin{itemize}
    \item Servers can \emph{restart} at any given point, but once they are \emph{all back up} and running the whole system should be operational again.
    
    \item Clients can connect to \emph{any} server, and perform any of these three operations:
    \begin{enumerate}
        \item Publish a message tagged by one or more topics.
        \item Subscribe to a list of topics.
        \item Get the last ten messages containing any of the topics he's subscribed to.
    \end{enumerate}
    
    \item Messages obtained by each client should reflect a \emph{coherent order} of all operations performed on the system by that same client or any other client.
\end{itemize}

% ---------------------------------------------------------------------------- %

% \paragraph{Assignment Assumptions}
% System servers are all known to each other. 

% \alberto{Briefly, what the system allows doing.}

% \alberto{Concretely, the requirements imposed by the assignment.}

% \alberto{Concretely, the assumptions allowed by the assignment.}

% \alberto{Our more refined view of the required functional system properties.}

% \alberto{Further, our additional assumptions.}


% Text.
% 
% \paragraph{Assumptions}

% Read-intensive workloads.

% \paragraph{Further assumptions and system properties}

% We assume that \emph{get} requests are submitted much more frequently than \emph{publish} requests.
% This affects the communication strategy between servers.

% We further refine some system properties that are not explicitly stated in the assignment:

% \begin{itemize}
    
%     \item When a client "gets" a list of messages and there is a message A on that list, then if the same client performs another "get", that same message A should be on this list as well, unless it was removed for there being ten messages more recent than A.   

%     \item All messages have to appear on the same order no matter the server from which the "get" was performed.

% \end{itemize}

% ---------------------------------------------------------------------------- %
